Postural changes mediate children's visual access to social information

### 2018-01-14 writeup-cogsci-copy/cogsci2018.Rmd
edits to text and citations

abstract: 
  "The ability to process social cues (e.g., eye gaze) is a critical component of children's early language and cognitive development. However, as children reach their first birthday, they begin to locomote themselves, exploring their visual environment in an entirely new way. How do these postural changes affect children's access to the social information relevant for word-learning? Here, explore this question by using head-mounted cameras to record infants' (8-16 months of age) egocentric visual perspective and by using computer-vision algorithms to detect the proportion of faces and hands in infants' environments.  We find that infant's postural changes largely mediate infants' accesses to both faces and hands, suggesting that these postural and locomotive developments facilitate infants' emerging linguistic and social capacities."

keywords:
    "social cognition; face-perception; infancy; locomotion; head-cameras"

@@

However, as children are learning their first words, their view of the world is also changing radically [@adolph2007]. Infant’s motor abilities improve dramatically near the end of the first year of life,  allowing them to locomote independently and to determine what they see. These motor changes have drastic consequences for what children see: crawling and walking infants have radically different views of the world. During spontaneous play in a laboratory playroom, toddlers are more likely to look at the floor while crawling than while walking (Franchak et al., 2011); walking vs. crawling infants tend to have full visual access to their environment and the people in it (Kretch, Franchak, and Adolph, 2014).

One possibility is that these motor improvements have strong developmental cascades, impacting children’s emerging social, cognitive, and linguistic abilities (Iverson, 2010). Indeed, these postural changes also impact how children interact with their mothers;  walking infants make different kinds of object-related bids for attention from their mothers than crawling infants, and tend to hear more action directed statements (e.g., “open it”).  (Karasik et al., 2014).  More directly, in an observational study, @walle2014 found that children who were able to walk had both higher receptive and productive vocabularies.  On their account, children’s ability to stand and independently locomote may fundamentally change their ability to access the social information (e.g., faces, gaze) relative to children who are still crawling and sitting. In other words, the ability to access more detailed social information may allow infants to learn words quicker and more efficiently, facilitating language growth.

Recent work has begun to use egocentric, head-mounted cameras to document the visual experiences of infants and children — which even for walking children are radically different than the adult perspective (and not easily predicted by our own adult intuitions; @franchak2011, @yoshida2008; @clerkin2017). Children’s views tend to be restricted and to be dominated by objects and hands much more than that of adults [@yoshida2008], and both computational and empirical work suggest that this restricted viewpoint may be more effective for learning what objects and their labels than the comparable adult perspective [@yurovsky2012, @bambach2017]. Further, recent work also suggests dramatic changes in the child’s perspective over the first two years of life, as views transition from primarily containing close up view of faces to capturing views of hands paired with the objects they are acting on [@fausey2017].

Here, we directly ask whether the postural changes that infants’ experience as they reach their first birthday change the availability of the social information relevant for word learning. To do so, we recorded the visual experience of a group of infants and children using head-mounted cameras during a brief laboratory free-play session; children’s posture and their orientation to their caregiver were also recorded from a third-person perspective and hand-annotated. We capitalize on recent improvements in computer vision algorithms (@citemtcnn, @openpose) that allow the automated detection of both faces and hands, and we analyze the frequency of both faces and hands in the child’s visual environment.

@@

## Face Detection
Measuring the availability of caregiver’s faces across development was an important goal of the study. We thus used a total of three face detection systems to explore this hypothesis and to avoid the cost of hand-annotating every frame. We first measure the performance of the most commonly-used and widely avlaialbe face detection algorithmsm (Viola-Jones). We use this as a benchmark for performance, as while it can achieve impressive accuracy in some situations, it is is notoriously bad at dealing with occluded faces (@citeAnthony2016). We next capitalized on recent improvements in computer vision, testing the performance of two other algorithms that extract face information in distinct ways.  Rsults for the evaluation of these three systems are shown in Table XX.

@@

### Face Detection Algorithms
The first face detection system made use of a series of Haar feature-based cascade classifiers (Viola-Jones, CITEXX) applied to each individual frame. This detector provided information about the presence of a face as well as its size and position.

The second algorithm was based on the work by (CITEXX) using multi-task cascaded convolutional neural networks (MTCNNs). The system was built using a novel cascaded CNN-based framework for joint detection and alignment, built to perform well in real-world environments where varying illuminations and occlusions are present. We used a Tensorflow implementation of this algorithm provided by (https://github.com/davidsandberg/facenet). Like Viola-Jones, this detector provided information about the presence of a face as well as it's size and position.

The third algorithm was a CNN-based pose detector that provided the locations of 18 body parts (ears, nose, wrists, etc) (CITEXX). The system uses a CNN for initial anatomical detection and subsequently applies part affinity fields (PAFs) for part association, producing a series of body part candidates. The candidates are then matched to a single individual and finally assembled into a pose. For the purposes of our investigation we only made use of the body parts relevant to the face (ears, eyes, nose). We operationalized face detections as any frames containing a nose and either the left or right eye; we operationalized hand detections as any frames containing any of the 21 points detected on either the left or the right hand, and we operationalized wrist detections any frames containing either the left or the right wrist.

@@

<!-- ```{r xtable, results="asis"} -->
<!-- tab3 <- xtable::xtable(summary(fullModelminsInt_Handd)$coef, digits=3, -->
<!--                       caption = "Results from GLMM model prediting the presence/absence of a hands (OpenPose-Hands) across the entire dataset.") -->

<!-- print(tab3, type="latex", comment = F, table.placement = "H") -->
<!-- ``` -->

>>>>

abstract: 
  "The ability to process social cues--including eye gaze--is a critical component of children's early language and cognitive development. However, as children reach their first birthday, they begin to locomote themselves, walking and exploring their visual environment in an entirely new way. How do these postural and locomotive changes affect children's access to the social information relevant for word-learning? Here, explore this question by using head-mounted cameras to record infants' (8-16 months of age) egocentric visual perspective and then using modern computer-vision algorithms to detect the proportion of faces and hands in infants' environments.  We find that infant's postural changes largely mediate infants' acess to both faces and hands, suggesting that these postural and locomotive developments facilitate infants' emerging linguistic and social capacities. Broadly, we suggest that the combined use of head-mounted cameras and the application of novel deep learning algorithms is a promising avenue for understanding the statistics of infants' visual and linguistic experience. "

keywords:
    "social cognition; face-perception; infancy; locomotion; head-cameras; deep learning"

@@

However, as children are learning their first words, their view of the world is also changing radically [@adolph2007]. Infant’s motor abilities improve dramatically near the end of the first year of life,  allowing them to locomote independently and to determine what they see. These motor changes have drastic consequences for what children see: crawling and walking infants have radically different views of the world. During spontaneous play in a laboratory playroom, toddlers are more likely to look at the floor while crawling than while walking [@franchak2011]; walking vs. crawling infants tend to have full visual access to their environment and the people in it [@kretch2014].

One possibility is that these motor improvements have strong developmental cascades, impacting children’s emerging social, cognitive, and linguistic abilities (Iverson, 2010). Indeed, these postural changes also impact how children interact with their mothers;  walking infants make different kinds of object-related bids for attention from their mothers than crawling infants, and tend to hear more action directed statements (e.g., “open it”) [@karasik2014].  More directly, in an observational study, @walle2014 found that children who were able to walk had both higher receptive and productive vocabularies.  On their account, children’s ability to stand and independently locomote may fundamentally change their ability to access the social information (e.g., faces, gaze) relative to children who are still crawling and sitting. In other words, the ability to access more detailed social information may allow infants to learn words quicker and more efficiently, facilitating language growth.

Recent work has begun to use egocentric, head-mounted cameras to document the visual experiences of infants and children — which even for walking children are radically different than the adult perspective (and not easily predicted by our own adult intuitions; @franchak2011, @yoshida2008; @clerkin2017). Children’s views tend to be restricted and to be dominated by objects and hands much more than that of adults [@yoshida2008], and both computational and empirical work suggest that this restricted viewpoint may be more effective for learning what objects and their labels than the comparable adult perspective [@yurovsky2012, @bambach2017]. Further, recent work also suggests dramatic changes in the child’s perspective over the first two years of life, as views transition from primarily containing close up view of faces to capturing views of hands paired with the objects they are acting on [@fausey2016].

Here, we directly ask whether the postural changes that infants’ experience as they reach their first birthday change the availability of the social information relevant for word learning. To do so, we recorded the visual experience of a group of infants and children using head-mounted cameras during a brief laboratory free-play session; children’s posture and their orientation to their caregiver were also recorded from a third-person perspective and hand-annotated. We capitalize on recent improvements in computer vision algorithms (@zhang2016, @cao2017realtime) that allow the automated detection of both faces and hands, and we analyze the frequency of both faces and hands in the child’s visual environment.

@@

## Face Detection
Measuring the availability of caregiver’s faces across development was an important goal of the study. We thus used a total of three face detection systems to explore this hypothesis and to avoid the cost of hand-annotating every frame. We first measure the performance of the most commonly-used and widely avlaialbe face detection algorithmsm (Viola-Jones). We use this as a benchmark for performance, as while it can achieve impressive accuracy in some situations, it is is notoriously bad at dealing with occluded faces (@scheirer2014perceptual). We next capitalized on recent improvements in computer vision, testing the performance of two other algorithms that extract face information in distinct ways.  Rsults for the evaluation of these three systems are shown in Table XX.

@@

### Face Detection Algorithms
The first face detection system made use of a series of Haar feature-based cascade classifiers (Viola-Jones, @viola2004robust) applied to each individual frame. This detector provided information about the presence of a face as well as its size and position.

The second algorithm was based on the work by @zhang2016 using multi-task cascaded convolutional neural networks (MTCNNs). The system was built using a novel cascaded CNN-based framework for joint detection and alignment, built to perform well in real-world environments where varying illuminations and occlusions are present. We used a Tensorflow implementation of this algorithm provided by (https://github.com/davidsandberg/facenet). Like Viola-Jones, this detector provided information about the presence of a face as well as it's size and position.

The third algorithm was a CNN-based pose detector that provided the locations of 18 body parts (ears, nose, wrists, etc) (@cao2017realtime, @simon2017hand, @wei2016cpm) avaliable at https://github.com/CMU-Perceptual-Computing-Lab/openpose. The system uses a CNN for initial anatomical detection and subsequently applies part affinity fields (PAFs) for part association, producing a series of body part candidates. The candidates are then matched to a single individual and finally assembled into a pose. For the purposes of our investigation we only made use of the body parts relevant to the face (ears, eyes, nose). We operationalized face detections as any frames containing a nose and either the left or right eye; we operationalized hand detections as any frames containing any of the 21 points detected on either the left or the right hand, and we operationalized wrist detections any frames containing either the left or the right wrist.

@@

# ```{r xtable, results="asis"}
# tab3 <- xtable::xtable(summary(fullModelminsInt_Handd)$coef, digits=3,
#                       caption = "Results from GLMM model prediting the presence/absence of hands (OpenPose-Hands) across the entire dataset.")
# 
# print(tab3, type="latex", comment = F, table.placement = "H")
# ```
